---
title: "Charles Clarke Assignment 1"
output:
  word_document: default
  pdf_document: default
  html_notebook: default
---


### Import libaries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(plotly)
library(ggplot2)
```







## Read data from CSV file

```{r}
MyData <- read.csv(file="data/bls-jobs-by-industry-category-1.csv", header=TRUE, sep=",")

```

```{r}
glimpse(MyData)

```

### Display summary of dataset
```{r}
summary(MyData)
```
##  structure of the dataset
```{r}
str(MyData)

```
### Print dataset data values.  Output 3 rows of data

```{r}
head(MyData,3)

```
```{r}

  
  
  
MyData  %>% 
group_by(Year)  %>% 
summarize(Total_Jobs = sum(Total.Jobs..private.and.government.))  %>% 
ggplot(aes(Year,Total_Jobs)) + geom_line() +
stat_smooth(method = "lm", color = "red", linetype="dotted") +
labs(title = "Total Jobs each Year",
    subtitle = "With a regression line fitting the Total Jobs")
```

###  Create new dataframe  and Load the cormat functions. The larger and darker the circle indicates strong correlations between the variables

```{r}

df_new <- data.frame(MyData$Total.Jobs..private.and.government. , MyData$Private.Sector..total., MyData$Mining..Logging.and.Construction, MyData$Year)
source("http://www.sthda.com/upload/rquery_cormat.r")
rquery.cormat(df_new,type="full")

```



### Plot Histogram of data.

```{r}
h<-hist(df_new$MyData.Total.Jobs..private.and.government., breaks=10, col="red", xlab="Total Jobs Private and Government", 
   main="Total private and government jobs")

```

### Plot bar chart of data


```{r}

barplot(df_new$MyData.Private.Sector..total.,
main = "Private Sector",
ylab = "Cost",
xlab = "Year",
names.arg = MyData$Year,
col = "darkred",
horiz = FALSE)



```
```{r}

barplot(df_new$MyData.Total.Jobs..private.and.government.,
main = "Total Private and Government Jobs",
ylab = "Cost",
xlab = "Year",
names.arg = MyData$Year,
col = "darkred",
horiz = FALSE)



```





### Print crosstab of data  (private and government)   by (year)

```{r}

### cross table
source("http://pcwww.liv.ac.uk/~william/R/crosstab.r")
report = crosstab(df_new,row.vars ="MyData.Year", type = "f")
print(report)
```






### Plot scatter plot of data

```{r}

plot(MyData$Total.Jobs..private.and.government,MyData$Year)
```














